Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-07-28 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Tweak some file paths

```{r echo = FALSE, fig.height = 4, fig.width = 6}
#library(data.table)
#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
                             start_year = 2010, end_year = 2010,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
alex_nldas <- alex_nldas %>%
        filter(storm_id == "Alex-2010")
alex_gauge$fips <- as.character(alex_gauge$fips)
compare_rain <- right_join(alex_nldas, alex_gauge, by = "fips")
#alex_nldas <- as.data.table(alex_nldas)
#alex_gauge <- as.data.table(alex_gauge)
#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
a
```

@@

```{r echo = FALSE, fig.height = 4, fig.width = 6}
library(weathermetrics)
#library(data.table)
#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
load("ike_gauge.RData")
ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
                                    "inches", "mm")
ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
                             start_year = 2008, end_year = 2008,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
ike_nldas <- ike_nldas %>%
        filter(storm_id == "Ike-2008")
ike_gauge$fips <- as.character(ike_gauge$fips)
compare_rain <- right_join(ike_nldas, ike_gauge, by = "fips")
#ike_nldas <- as.data.table(alex_nldas)
#ike_gauge <- as.data.table(alex_gauge)
#compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
b <- ggplot(compare_rain1, aes(x = tot_precip, y = rain_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
b
```

@@

```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
compare_rain %>% group_by(fips) %>%
        summarize(ave_rain_mm = mean(rain_mm),
                  ave_tot_precip = mean(tot_precip),
                  n = n(),
                  within_500_km = first(storm_dist) <= 500) %>%
        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
                   size = n, color = within_500_km)) + 
        geom_hline(yintercept = 75, color = "gray") + 
        geom_vline(xintercept = 75, color = "gray") + 
        geom_point(alpha = 0.5) + 
        theme_bw() +
        xlim(c(10, 225)) + ylim(c(10, 225)) + 
        geom_abline(intercept = 0, slope = 1) + 
        xlab("Rainfall based on NLDAS-2 data (mm)") + 
        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
        scale_color_manual(values = c("blue", "black"), guide = "none") 
```

@@

```{r}
#frances_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_gauge.csv", header = TRUE)
#save(frances_gauge, file = "writing/DraftExposurePaper/frances_gauge.RData")
#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_gauge.RData")
frances_gauge$precip_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
frances_nldas <- county_rain(counties = unique(frances_gauge$fips),
                             start_year = 2004, end_year = 2004,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
frances_nldas <- frances_nldas %>%
        filter(storm_id == "Frances-2004")
frances_gauge$fips <- as.character(frances_gauge$fips)
compare_rain <- right_join(frances_nldas, frances_gauge, by = "fips")
b <- ggplot(compare_rain, aes(x = tot_precip, y = precip_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
b
```

@@

```{r}
#Ike-2008 wind
#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
#save(ike_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
```

@@

```{r}
frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", na.strings = 999,
                         header = TRUE)
save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_wind.RData")
```

@@

```{r}
#devtools::install_github("ropenscilabs/rnoaa")
library(rnoaa)
#devtools::install_github("leighseverson/countyweather")
library(countyweather)
options("noaakey" = Sys.getenv("noaakey"))
#Alex-2010
alex_fips <- alex_nldas$fips
alex_fips <- unique(alex_fips)
#Ike-2008
ike_fips <- ike_nldas$fips
ike_fips <- unique(ike_fips)
```

@@

```{r}
#Find the top 15 FIPS by number of reporting stations during Ike-2008 (Warning: Takes a long time)
#my_fips <- ike_fips[c(1:6, 10, 13:16, 18, 20:25, 28:29, 31, 33:34, 36:40, 42, 44:45)]
#my_stations <- NA
#ike_table <- expand.grid(my_fips, my_stations)
#colnames(ike_table) <- c("fips", "station")
#for(fips in my_fips){
        #ike_weather <- weather_fips(fips = fips,
                                    #date_min = "2008-09-01",
                                    #date_max = "2008-09-14",
                                    #var = "PRCP")
        #ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
#} 
#Ike FIPS without data: 7, 8, 9, 11, 12, 17, 19, 26, 27, 30, 32,35, 41, 43, 
#ike_table <- ike_table %>%
        #arrange(desc(station)) %>%
        #slice(1:15)
#ike_top <- ike_table$fips
#Find the top 15 FIPS by number of reporting stations for Alex-2010:
#my_fips <- alex_fips[c(1:2, 4:6, 8, 10, 12:14, 16:18)]
#my_stations <- NA
#alex_table <- expand.grid(my_fips, my_stations)
#colnames(alex_table) <- c("fips", "station")
#for(fips in my_fips){
#        alex_weather <- weather_fips(fips = fips,
#                                    date_min = "2008-09-01",
#                                    date_max = "2008-09-14",
#                                    var = "PRCP")
#        alex_table[alex_table$fips == fips, "station"] <- max(alex_weather$daily_weather$prcp_reporting)    
#} 
#Alex FIPS without data: 3, 7, 9, 11, 15, 19   
#alex_table <- alex_table %>%
#        arrange(desc(station)) %>%
#        slice(1:15)
#alex_top <- alex_table$fips
#fips_table <- rbind(alex_table, ike_table)
#fips_table <- fips_table %>%
#        arrange(desc(station)) %>%
#        slice(1:15)
#all_fips <- fips_table$fips
#all_fips <- unique(all_fips)
```

As of right now, I am not going to do Frances (there are 184 unique fips).

Create a timeseries for the top 15 FIPS by reporting stations for Alex and Ike:

```{r}
data("closest_dist")
data(rain)
county_timeseries(fips = ike_top[1:2], 
                  percent_coverage = 0, 
                  date_min = "2008-09-01", 
                  date_max = "2010-09-14", var = "PRCP", 
                  out_directory = "~/tmp/ike_ex")
ike_1 <- readRDS("~/tmp/ike_ex/12087.rds")
keys_cd <- subset(closest_dist, closest_dist$storm_id == "Ike-2008" & closest_dist$fips == "12087")
keys_cd <- keys_cd$closest_date
ike_1 <- ike_1 %>%
        select(date, prcp) %>%
        filter(date %in% seq(from = as.Date("2008-09-07"), to = as.Date("2008-09-10"), by = 1))
ike_rain <- subset(rain, rain$storm_id == "Ike-2008" & rain$fips == "12087")
ike_sum <- sum(ike_1$prcp)
ike_nldas_sum <- subset(ike_nldas, ike_nldas$fips == "12087")
ike_nldas_sum <- ike_nldas_sum$tot_precip
```

[JMF-- It looks like precipitation measurements from `countyweather` are already in mm]

>>>>

```{r echo = FALSE, fig.height = 4, fig.width = 6}
#library(data.table)
#alex_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_Alex_Rain.csv", header = TRUE)
#save(alex_gauge, file = "writing/DraftExposurePaper/alex_gauge.RData")
# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/alex_gauge.Rdata")
load("alex_gauge.Rdata")
alex_nldas <- county_rain(counties = unique(alex_gauge$fips),
                             start_year = 2010, end_year = 2010,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
alex_nldas <- alex_nldas %>%
        filter(storm_id == "Alex-2010")
alex_gauge$fips <- as.character(alex_gauge$fips)
compare_rain <- right_join(alex_nldas, alex_gauge, by = "fips")
#alex_nldas <- as.data.table(alex_nldas)
#alex_gauge <- as.data.table(alex_gauge)
#compare_rain <- left_join(alex_nldas, alex_gauge, by = "fips")
a <- ggplot(compare_rain, aes(x = tot_precip, y = rain_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
a
```

@@

```{r echo = FALSE, fig.height = 4, fig.width = 6}
library(weathermetrics)
#library(data.table)
#ike_gauge <- read.csv("/Users/joshuaferreri/Documents/ike_rain.csv", header = TRUE)
#save(ike_gauge, file = "writing/DraftExposurePaper/ike_gauge.RData")
#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_gauge.RData")
load("ike_gauge.RData")
ike_gauge$rain_mm <- convert_precip(ike_gauge$rain_in, 
                                    "inches", "mm")
ike_nldas <- county_rain(counties = unique(ike_gauge$fips),
                             start_year = 2008, end_year = 2008,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
ike_nldas <- ike_nldas %>%
        filter(storm_id == "Ike-2008")
ike_gauge$fips <- as.character(ike_gauge$fips)
compare_rain1 <- right_join(ike_nldas, ike_gauge, by = "fips")
#ike_nldas <- as.data.table(alex_nldas)
#ike_gauge <- as.data.table(alex_gauge)
#compare_rain1 <- left_join(alex_nldas, alex_gauge, by = "fips")
b <- ggplot(compare_rain1, aes(x = tot_precip, y = rain_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
b
```

@@

```{r echo = FALSE, fig.width = 5.5, fig.height = 4}
compare_rain1 %>% group_by(fips) %>%
        summarize(ave_rain_mm = mean(rain_mm),
                  ave_tot_precip = mean(tot_precip),
                  n = n(),
                  within_500_km = first(storm_dist) <= 500) %>%
        ggplot(aes(x = ave_tot_precip, y = ave_rain_mm, 
                   size = n, color = within_500_km)) + 
        geom_hline(yintercept = 75, color = "gray") + 
        geom_vline(xintercept = 75, color = "gray") + 
        geom_point(alpha = 0.5) + 
        theme_bw() +
        xlim(c(10, 225)) + ylim(c(10, 225)) + 
        geom_abline(intercept = 0, slope = 1) + 
        xlab("Rainfall based on NLDAS-2 data (mm)") + 
        ylab("Rainfall reported in Monthly \nWeather Report (mm)") + 
        scale_color_manual(values = c("blue", "black"), guide = "none") 
```

@@

```{r}
#frances_gauge <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_gauge.csv", header = TRUE)
#save(frances_gauge, file = "writing/DraftExposurePaper/frances_gauge.RData")
#load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/frances_gauge.RData")
load("frances_gauge.RData")
frances_gauge$precip_mm <- convert_precip(frances_gauge$precip_in, "inches", "mm")
frances_nldas <- county_rain(counties = unique(frances_gauge$fips),
                             start_year = 2004, end_year = 2004,
                             rain_limit = 0, dist_limit = 1500,
                             days_included = c(-2, -1, 0, 1))
frances_nldas <- frances_nldas %>%
        filter(storm_id == "Frances-2004")
frances_gauge$fips <- as.character(frances_gauge$fips)
compare_rain <- right_join(frances_nldas, frances_gauge, by = "fips")
b <- ggplot(compare_rain, aes(x = tot_precip, y = precip_mm,
                              color = fips)) +
        geom_point() + 
        theme_bw()
b
```

@@

```{r}
#Ike-2008 wind
#ike_wind <- read.csv("/Users/joshuaferreri/Documents/ike_wind.csv", na.strings = 999, header = TRUE)
#save(ike_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
# load("~/Documents/hurricaneexposure/writing/DraftExposurePaper/ike_wind.RData")
load("ike_wind.RData")
```

@@

```{r eval = FALSE}
frances_wind <- read.csv("/Users/joshuaferreri/Documents/NWR_frances_wind.csv", 
                         na.strings = 999,
                         header = TRUE)
save(frances_wind, file = "writing/DraftExposurePaper/frances_wind.RData")
```

```{r}
load("frances_wind.RData")
```

@@

```{r eval = FALSE}
#devtools::install_github("ropenscilabs/rnoaa")
library(rnoaa)
#devtools::install_github("leighseverson/countyweather")
library(countyweather)
options("noaakey" = Sys.getenv("noaakey"))
#Alex-2010
alex_fips <- alex_nldas$fips
alex_fips <- unique(alex_fips)
#alex_weather <- weather_fips(fips = alex_fips,
                             #date_min = "2010-06-25",
                             #date_max = "2010-07-02",
                             #var = "PRCP") # Returns warning: "sorry, no data found"
                                            # and problemin station_ids$meta . . .
#Ike-2008
ike_fips <- ike_nldas$fips
ike_fips <- unique(ike_fips)
#ike_weather <- county_timeseries(fips = ike_fips,
                                 #date_min = "2008-09-01",
                                 #date_max = "2008-09-14",
                                 #out_directory = paste0("/Users/joshuaferreri",
                                                # "/Documents/ike_timeseries"),
                                 #out_type = ".csv")
ike_weather <- weather_fips(fips = ike_fips[1],
                             date_min = "2008-09-01",
                             date_max = "2008-09-14",
                             var = "PRCP") 
```

@@

```{r eval = FALSE}
my_fips <- ike_fips[1:3]
my_stations <- NA
test <- expand.grid(my_fips, my_stations)
colnames(test) <- c("fips", "station")
for(fips in my_fips){
        ike_weather <- weather_fips(fips = fips,
                                    date_min = "2008-09-01",
                                    date_max = "2008-09-14",
                                    var = "PRCP")
        test[test$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)    
} #This works fine (small subset of code below)
ike_table <- expand.grid(ike_fips, my_stations)
colnames(ike_table) <- c("fips", "station")
#for(fips in ike_fips){
        ike_weather <- weather_fips(fips = fips,
                                    date_min = "2008-09-01",
                                    date_max = "2008-09-14",
                                    var = "PRCP")
        ike_table[ike_table$fips == fips, "station"] <- max(ike_weather$daily_weather$prcp_reporting)
} #getting error
```
